### 2015-01-03 paper/BMC_Bioinformatics_2015/Fig_MelnikovDiffPrefs/Fig_MelnikovDiffPrefs.tex
Added manuscript draft for BMC Bioinformatics

\documentclass[12pt,letterpaper]{article}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{rotating}
\usepackage[export]{adjustbox}
\pagestyle{empty}
\usepackage[paperheight=5.5in,paperwidth=6.4in,top=0.05in,bottom=0.05in,right=0.05in,left=0.05in]{geometry}

\begin{document}
\begin{figure}

{\bf \Large A} \\
\includegraphics[clip, trim=0 0 0 0.04in,valign=t,width=5.2in]{../../../examples/Melnikov_et_al_Tn5/infer_diffprefs_on_simulated_data/inferred_diffprefs_logoplot.pdf}
\vspace{0.15in}

{\bf \Large B} \\
\includegraphics[clip, trim=0.12in 0 0 0.05in,valign=t,width=6.3in]{../../../examples/Melnikov_et_al_Tn5/infer_diffprefs_on_simulated_data/correlations.pdf}

\end{figure}
\end{document}