### 2015-01-03 paper/BMC_Bioinformatics_2015/Fig_MelnikovPrefs/Fig_MelnikovPrefs.tex

\documentclass[12pt,letterpaper]{article}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{rotating}
\usepackage[export]{adjustbox}
\pagestyle{empty}
\usepackage[paperheight=4.1in,paperwidth=6.73in,top=0.05in,bottom=0.05in,right=0.05in,left=0.05in]{geometry}

\begin{document}
\begin{figure}
\begin{minipage}[t][][t]{5in}
{\bf \Large A} \\
\includegraphics[clip, trim=0 0 0 0.04in,valign=t,width=5in]{../../../examples/Melnikov_et_al_Tn5/prefs_logoplot.pdf}
\end{minipage}
\hspace{0.05in}
\begin{minipage}[t][][t]{1.5in}
{\bf \Large B} \\
\vspace{-0.2in}
\includegraphics[valign=t,width=1.5in]{../../../examples/Melnikov_et_al_Tn5/replicate_pref_corr.pdf}
\end{minipage}
\end{figure}
\end{document}