### 2015-01-03 paper/BMC_Bioinformatics_2015/Fig_ThyagarajanPrefs/Fig_ThyagarajanPrefs.tex

\documentclass[12pt,letterpaper]{article}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{rotating}
\usepackage[export]{adjustbox}
\pagestyle{empty}
\usepackage[paperheight=7.6in,paperwidth=6.3in,top=0.05in,bottom=0.05in,right=0.05in,left=0.05in]{geometry}

\begin{document}
\begin{figure}

{\bf \Large A} \\
\includegraphics[clip, trim=0 0 0 0.04in,valign=t,width=6.3in]{../../../examples/ThyagarajanBloom2014_WSN_HA/prefs_logoplot.pdf}
\vspace{0.07in}

{\bf \Large B} \\
\vspace{-0.2in}
\includegraphics[valign=t,width=1.5in]{../../../examples/ThyagarajanBloom2014_WSN_HA/correlations/corr_1_vs_2.pdf}\hspace{0.2in}
\includegraphics[valign=t,width=1.5in]{../../../examples/ThyagarajanBloom2014_WSN_HA/correlations/corr_1_vs_3.pdf}\hspace{0.2in}
\includegraphics[valign=t,width=1.5in]{../../../examples/ThyagarajanBloom2014_WSN_HA/correlations/corr_2_vs_3.pdf}

\end{figure}
\end{document}